---
title: "Optimal allocations: Case study / Simulations"
# author: "Marta Bofill Roig"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: hide 
    theme: flatly  
    includes:
      after_body: footer.html
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```


<!-- <a href="https://github.com/MartaBofillRoig" class="github-corner" aria-label="View source on GitHub"><svg width="80" height="80" viewBox="0 0 250 250" style="fill:#151513; color:#fff; position: absolute; top: 0; border: 0; right: 0;" aria-hidden="true"><path d="M0,0 L115,115 L130,115 L142,142 L250,250 L250,0 Z"></path><path d="M128.3,109.0 C113.8,99.7 119.0,89.6 119.0,89.6 C122.0,82.7 120.5,78.6 120.5,78.6 C119.2,72.0 123.4,76.3 123.4,76.3 C127.3,80.9 125.5,87.3 125.5,87.3 C122.9,97.6 130.6,101.9 134.4,103.2" fill="currentColor" style="transform-origin: 130px 106px;" class="octo-arm"></path><path d="M115.0,115.0 C114.9,115.1 118.7,116.5 119.8,115.4 L133.7,101.6 C136.9,99.2 139.9,98.4 142.2,98.6 C133.8,88.0 127.5,74.4 143.8,58.0 C148.5,53.4 154.0,51.2 159.7,51.0 C160.3,49.4 163.2,43.6 171.4,40.1 C171.4,40.1 176.1,42.5 178.8,56.2 C183.1,58.6 187.2,61.8 190.9,65.4 C194.5,69.0 197.7,73.2 200.1,77.6 C213.8,80.2 216.3,84.9 216.3,84.9 C212.7,93.1 206.9,96.0 205.4,96.6 C205.1,102.4 203.0,107.8 198.3,112.5 C181.9,128.9 168.3,122.5 157.7,114.1 C157.9,116.9 156.7,120.9 152.7,124.9 L141.0,136.5 C139.8,137.7 141.6,141.9 141.8,141.8 Z" fill="currentColor" class="octo-body"></path></svg></a><style>.github-corner:hover .octo-arm{animation:octocat-wave 560ms ease-in-out}@keyframes octocat-wave{0%,100%{transform:rotate(0)}20%,60%{transform:rotate(-25deg)}40%,80%{transform:rotate(10deg)}}@media (max-width:500px){.github-corner:hover .octo-arm{animation:none}.github-corner .octo-arm{animation:octocat-wave 560ms ease-in-out}}</style> -->

# Preface 

Load R packages and functions

```{r}
library(tidyverse)
# install.packages("plyr")
library(plyr)
# devtools::install_github("pavlakrotka/NCC@v1.0")
library(NCC)
source("C:/Users/mbofi/Dropbox/CeMSIIS/GitHub/Allocation/case-study/aux_functions.R")
```

# Case Study


We illustrate the optimal allocations in platform trials by means of a  phase  II  placebo-controlled  trial  in primary  hypercholesterolemia. 

```{r, echo=F}
# sample sizes original study
# n_control = 31
# n_arm1 = 31
# n_arm2 = 30
# N = n_control + n_arm1 + n_arm2 
# N1=n_arm1

# Sample sizes for the simulation 
N = 80
# N1 = round(N/3)
# N2 = round(2*(N-N1)/3)
# c(N1,N2,N-N1-N2)
```

In the original study, $N=92$ patients were randomised following 1:1:1. 

```{r} 
# means
mean_control = 17.3/3.5
mean_arm1 = 66.2/3.5
mean_arm2 = 72.3/3.5
```


In what follows, we simulated the trial using the estimated mean in the control arm in the original study using three allocation strategies -namely, equal allocation (1:...:1), square root of $k$ (1:...:$\sqrt(k)$), and the proposed optimal allocations-, and according to three different trial designs:

  (1) Design with one period only (that is, multi-arm design)
  (2) Design with two periods (arm 2 starts later, but arms 1 and 2 finish at the same time)
  (3) Design with three periods (arm 2 starts later and finishes after arm 1 does)

  
For comparative purposes, in this case study, we suppose total sample size of $N=80$ and smaller effect sizes.  

## Design 1: multi-arm design

In this case, we consider a design with one period only. The scheme of the trial over time is:

```{r, fig.align="center", fig.width=4, fig.height=4, fig.cap="Figure: Design 1: multi-arm design."}
db1_one = sim_designs(r1=1,r2=0,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="one")
db1_sqrt = sim_designs(r1=1,r2=0,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="sqrt")
db1_opt = sim_designs(r1=1,r2=0,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="opt")
plot_trial(db1_one$data$treatment) 
```

Distribution of sample sizes per arm and periods
```{r}
# sample sizes
db1_one$ss
db1_sqrt$ss
db1_opt$ss

db1_one_ss <- data.frame(arms=c("A1","A2","C"),db1_one$ss, c(sum(db1_one$ss[1,]),sum(db1_one$ss[2,]),sum(db1_one$ss[3,])))
db1_sqrt_ss <- data.frame(arms=c("A1","A2","C"), db1_sqrt$ss, c(sum(db1_sqrt$ss[1,]),sum(db1_sqrt$ss[2,]),sum(db1_sqrt$ss[3,])))
db1_opt_ss <- data.frame(arms=c("A1","A2","C"), db1_opt$ss, c(sum(db1_opt$ss[1,]),sum(db1_opt$ss[2,]),sum(db1_opt$ss[3,])))
```


The sample sizes per arm and period according to the allocation strategies are the following:
```{r}
knitr::kable(db1_one_ss, format = "markdown", caption = c("Sample sizes per period and arm (1:1)"), col.names = c("Arms", "Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db1_sqrt_ss, format = "markdown", caption = c("Sample sizes per period and arm (sqrt(k)-rule)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db1_opt_ss, format = "markdown", caption = c("Sample sizes per period and arm (optimal allocations)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
```


Comparing groups when using 1:1 allocation
```{r}
res1_one = do.call(rbind.data.frame, models_cc(data = db1_one$data) )
knitr::kable(res1_one, format = "markdown") 
```
Comparing groups when using $\sqrt(k)$-allocation (and thus optimal allocations)
```{r}
res1_opt = do.call(rbind.data.frame, models_cc(data = db1_opt$data) )
knitr::kable(res1_opt, format = "markdown")
```

<!-- ```{r} -->
<!-- res1_sqrt = do.call(rbind.data.frame, models_cc(data = db1_sqrt$data) ) -->
<!-- knitr::kable(res1_sqrt, format = "markdown") -->
<!-- ``` -->



## Design 2: two-period design

```{r}
N = 80
N1 = round(N/4)
N2 = round(N-N1)
c(N1,N2,N-N1-N2)
```


In this case, we consider a design with two periods. The scheme of the trial over time is:
```{r, fig.align="center", fig.width=4, fig.height=4, fig.cap="Figure: Design 2: two-period design."} 
db2_one=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="one")
db2_sqrt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="sqrt")
db2_opt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="opt")
plot_trial(db2_one$data$treatment) 
```

```{r}
# sample sizes
db2_one$ss
db2_sqrt$ss
db2_opt$ss

db2_one_ss <- data.frame(arms=c("A1","A2","C"),db2_one$ss, c(sum(db2_one$ss[1,]),sum(db2_one$ss[2,]),sum(db2_one$ss[3,])))
db2_sqrt_ss <- data.frame(arms=c("A1","A2","C"), db2_sqrt$ss, c(sum(db2_sqrt$ss[1,]),sum(db2_sqrt$ss[2,]),sum(db2_sqrt$ss[3,])))
db2_opt_ss <- data.frame(arms=c("A1","A2","C"), db2_opt$ss, c(sum(db2_opt$ss[1,]),sum(db2_opt$ss[2,]),sum(db2_opt$ss[3,])))
```


The sample sizes per arm and period according to the allocation strategies are the following:
```{r}
knitr::kable(db2_one_ss, format = "markdown", caption = c("Sample sizes per period and arm (1:1)"), col.names = c("Arms", "Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db2_sqrt_ss, format = "markdown", caption = c("Sample sizes per period and arm (sqrt(k)-rule)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db2_opt_ss, format = "markdown", caption = c("Sample sizes per period and arm (optimal allocations)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
```


Comparing groups when using 1:1 allocation
```{r}
res2_one = do.call(rbind.data.frame, models_cc(data = db2_one$data) )
knitr::kable(res2_one, format = "markdown") 
```
Comparing groups when using $\sqrt(k)$-allocation
```{r}
res2_sqrt = do.call(rbind.data.frame, models_cc(data = db2_sqrt$data) )
knitr::kable(res2_sqrt, format = "markdown")
```
Comparing groups when using the optimal allocations
```{r}
res2_opt = do.call(rbind.data.frame, models_cc(data = db2_opt$data) )
knitr::kable(res2_opt, format = "markdown")
```

## Design 3: three-period design

Suppose now a design with three periods with $N_1=31$ and consider two situations for $N_2$, say $N_2=N-N_1$ and $N_2= N_1/2$. 

### Trial with equal allocation rates for periods 1 and 3

Suppose now that the size of the periods are:
```{r}
N1 = round(N/3)
N2 = round(N-2*N1)
c(N, N1, N2, N-N1-N2)
```

Note that in this case the duration of periods 1 and 3 is the same, leading to a symmetrical trial. 
Below we illustrate the scheme of the trial over time. 

```{r, fig.align="center", fig.width=4, fig.height=4, fig.cap="Design 3: three-period design (r1=r3)."} 

db3_one=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="one")
db3_sqrt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="sqrt")
db3_opt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="opt")

plot_trial(db3_opt$data$treatment) 

# sample sizes
db3_one$ss
db3_sqrt$ss
db3_opt$ss

db3_one_ss <- data.frame(arms=c("A1","A2","C"),db3_one$ss, c(sum(db3_one$ss[1,]),sum(db3_one$ss[2,]),sum(db3_one$ss[3,])))
db3_sqrt_ss <- data.frame(arms=c("A1","A2","C"), db3_sqrt$ss, c(sum(db3_sqrt$ss[1,]),sum(db3_sqrt$ss[2,]),sum(db3_sqrt$ss[3,])))
db3_opt_ss <- data.frame(arms=c("A1","A2","C"), db3_opt$ss, c(sum(db3_opt$ss[1,]),sum(db3_opt$ss[2,]),sum(db3_opt$ss[3,])))
```

The sample sizes per arm and period according to the allocation strategies are the following:
```{r}
knitr::kable(db3_one_ss, format = "markdown", caption = c("Sample sizes per period and arm (1:1)"), col.names = c("Arms", "Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db3_sqrt_ss, format = "markdown", caption = c("Sample sizes per period and arm (sqrt(k)-rule)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db3_opt_ss, format = "markdown", caption = c("Sample sizes per period and arm (optimal allocations)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
```


Comparing groups when using 1:1 allocation
```{r}
res3_one = do.call(rbind.data.frame, models_cc(data = db3_one$data) )
knitr::kable(res3_one, format = "markdown") 
```
Comparing groups when using $\sqrt(k)$-allocation
```{r}
res3_sqrt = do.call(rbind.data.frame, models_cc(data = db3_sqrt$data) )
knitr::kable(res3_sqrt, format = "markdown")
```
Comparing groups when using the optimal allocations
```{r}
res3_opt = do.call(rbind.data.frame, models_cc(data = db3_opt$data) )
knitr::kable(res3_opt, format = "markdown")
```

### Trial with unequal allocation rates for for periods 1 and 3

Suppose now that the size of the periods are:
```{r}
# N = 80
N1 = round(N/3)
N2 = round(2*(N-N1)/3)
c(N1,N2,N-N1-N2) 
```


```{r, fig.align="center", fig.width=4, fig.height=4, fig.cap="Design 3: three-period design (r1<r3)."}
db3_one=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="one")
db3_sqrt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="sqrt")
db3_opt=sim_designs(r1=N1/N,r2=N2/N,mu0=mean_control,mu1=mean_arm1,mu2=mean_arm2,N=N,alloc="opt")

plot_trial(db3_opt$data$treatment) 

# sample sizes
db3_one$ss
db3_sqrt$ss
db3_opt$ss

db3_one_ss <- data.frame(arms=c("A1","A2","C"),db3_one$ss, c(sum(db3_one$ss[1,]),sum(db3_one$ss[2,]),sum(db3_one$ss[3,])))
db3_sqrt_ss <- data.frame(arms=c("A1","A2","C"), db3_sqrt$ss, c(sum(db3_sqrt$ss[1,]),sum(db3_sqrt$ss[2,]),sum(db3_sqrt$ss[3,])))
db3_opt_ss <- data.frame(arms=c("A1","A2","C"), db3_opt$ss, c(sum(db3_opt$ss[1,]),sum(db3_opt$ss[2,]),sum(db3_opt$ss[3,])))
```

The sample sizes per arm and period according to the allocation strategies are the following:
```{r}
knitr::kable(db3_one_ss, format = "markdown", caption = c("Sample sizes per period and arm (1:1)"), col.names = c("Arms", "Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db3_sqrt_ss, format = "markdown", caption = c("Sample sizes per period and arm (sqrt(k)-rule)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
knitr::kable(db3_opt_ss, format = "markdown", caption = c("Sample sizes per period and arm (optimal allocations)"), col.names = c("Arms","Period 1","Period 2","Period 3", "Total per arm"))
```


Comparing groups when using 1:1 allocation
```{r}
res3_one = do.call(rbind.data.frame, models_cc(data = db3_one$data) )
knitr::kable(res3_one, format = "markdown") 
```
Comparing groups when using $\sqrt(k)$-allocation
```{r}
res3_sqrt = do.call(rbind.data.frame, models_cc(data = db3_sqrt$data) )
knitr::kable(res3_sqrt, format = "markdown")
```
Comparing groups when using the optimal allocations
```{r}
res3_opt = do.call(rbind.data.frame, models_cc(data = db3_opt$data) )
knitr::kable(res3_opt, format = "markdown")
```

# Simulations

```{r}
load("C:/Users/mbofi/Dropbox/CeMSIIS/GitHub/Allocation/case-study/results/simstudy_results.RData")
df_res$design = ifelse(as.numeric(df_res$r1)+as.numeric(df_res$r2)==1,"2-period", "3-period")
```


To compare power and type 1 error of the different designs, we undertake a simulation study to evaluate the performance when using 1:1 allocations. For comparative purposes, we also consider a total sample size for the trial equal to XX 

```{r}
res_report_H1 <- df_res %>% filter(H0=="FALSE") %>% select(minrt,rt_a1,rt_a2,r1,r2,alloc,design)
knitr::kable(res_report_H1, format = "markdown", caption = c("Power comparisons"), col.names=c("Min Power", "Power A1",	"Power A2",	"r1",	"r2",	"Allocation",	"Design"))
```

```{r}
res_report_H0 <- df_res %>% filter(H0=="TRUE") %>% select(minrt,rt_a1,rt_a2,r1,r2,alloc,design)
knitr::kable(res_report_H0, format = "markdown", caption = c("Type 1 error rate"), col.names=c("Min T1E", "T1E A1",	"T1E A2",	"r1",	"r2",	"Allocation",	"Design"))
```


